%!TEX root = ../thesis.tex

\chapter{Introduction}
\label{chap:intro}

%Ideas->Describing general picture of melodic analysis for description and characterization of IAM. The one we have used in several presentations.
%Ideas-> When you talk about MIR first say what kinds of varied tasks. And then similarity at diff level and pattern recognition has been a focus since a long time! Talk about glass ceiling in similarities, (ref to Joan's thesis). This maybe because similarities has many variablesm and cultural influence is quite recognized one in that. 
%Ideas-> Our ideas behinf using characteristic phrase is like cover song similarity. There is no comparison between two items for similarity. Its from ones mental abstractino or image. See joan's work to get more insights on how has he presented it.
%Ideas-> read some more thesis on this topic and read intros from PR IITB, they are really good points 
%
%
%Overall what all things
%1) What are in generals the problems that world is facing related with digital music
%2) What if machines can understand music like humans do, what can they do
%
%3) MIR dedicated to this kind of research
%4) Varikous topics, variety of problems, basic research, applied reearch, interdesciplinary, mixture of approaches borrowed from other domains, 
%
%
%Joan's approach: Straight starting from task at hand, motivating need for it --> MIR-->

%
%Basically I want to say that data and information is increasing at a fast pace. We ICT tech that can understand, interpret and make sense of digital data like humans do, so that they can make our interaction with the data and information better. So that we can utilize the data and information better in our lives. Many factors play role here, one of them is the cultural and social context. So bsically those factors should also be taken into account. Once these systems learn how to make sense of data, not only will we be able to consume that better but doing this exercise of making machines make sense of the data can also increase our understanding of the concepts surrounding us. This can also enable novel ways od doing stuff, far suprior than what human do for now. 
%
%

%Technologies that can process digital data to understand and interpret it like humans do have becomes nearly a necessity in modern world. Such technology can not only perform tasks that humans already do but can also pave way for novel ways to interact with the data. During the process of developing such technologies we learn and undertand more about the phenomenon in the world. 






\section{Motivation, Context and Relevance}
\label{sec:intro_motivation_context_relevance}

%Repeating structures are important information units in data such as text, DNA sequences, images, videos, speech and music~\citep{Buhler2002b,Herley2006}. Patterns are exploited in a variety of ways, ranging from signal level tasks such as data-compression~\citep{Atallah1999} to more cognitively complex tasks such as analyzing an art work~\citep{van2010texton}. In music domain, identification of repeating structures in a musical piece is fundamental to its analysis, understanding and interpretation~\citep{Cook1987,Lerdahl1983}. 

\subsection{\titlecap{\glsentrylong{mir}}}
\label{sec:intro_motivation_mir}

\Gls{mir} is a growing interdisciplinary research field that primarily addresses topics involved in understanding and modeling of music using information processing methodologies~\citep{roadmap_mir}. In particular, it aims to advance our knowledge in representing, understanding, describing, retrieving and organizing music related data, which opens up wealth of possibilities to develop novel ways to interact with music~\citep{casey2008content,orio2006music,burgoyne2016music}. As mentioned \gls{mir} is interdisciplinary field, which stands at intersection of well established disciplines such as signal processing, pattern recognition, musicology, psychoacoustics, music perception and cognition, information science, and computer science (machine-learning). \TODO{remark that we are mainly talking about content based MIR}

A significant effort in \gls{mir} is dedicated towards automatic description and characterization of music content to extract musically and semantically relevant information. Computational approaches in \gls{mir} for tasks such as melody extraction~\citep{salamon:phd:13}, chord estimation~\citep{mcvicar2014automatic}, music key estimation~\citep{peeters2006chroma}, motif discovery~\citep{collins2011improved,Conklin2010}, music structure analysis~\citep{paulus2010state,serra2012unsupervised}, music similarity~\citep{joan_thesis}, genre classification~\cite{aucouturier2003representing} and emotion recognition~\citep{kim2010music}, tempo and beat tracking~\citep{gouyon2005review,scheirer1998tempo} describe and characterize music pieces in terms of different musical aspects such as melody, rhythm, harmony, structure and emotion. However, the definition, interpretation and relevance of these musical aspects are not universal and vary significantly across different music traditions,  and personal, cultural and social contexts. \TODO{You can add Xavier's sentence on his interpretation of description of MIR systems} Failing to account many of them, a significant number of existing computational approaches in \gls{mir} might be hitting the so-called ``glass ceiling"~\citep{pachet2004improving,casey2008content}. This phenomenon is fairly evident from the slowing rate of improvement of the approaches in \gls{mir} over years, as seen in the results of MIREX\footnote{http://www.music-ir.org/mirex/wiki/MIREX\_HOME} evaluations. There exists a semantic-gap between the automatically extracted music descriptors from audio signals and the high level music concepts that humans relate to~\citep{celma2006foafing,casey2008content}. There is a need to bridge this gap and to take a broader perspective to describe music by also considering the cultural, social and user context into account~\citep{roadmap_mir}. \TODO{highlight lack of knowledge infusion from top dwn. Most approaches work on audio based on features and try to describe higher level concepts}.

\TODO{Bob sturm stuff, rephrase}. In this context, it is worth mentioning the recent debates on whether machines are really understanding or ``making sense" of music or are just producing good classification numbers in evaluations \TODO{ref: bob sturm horse paper}. This skepticism of our interpretation of several \gls{mir} systems become evident when the trained methods for a number of \gls{mir} tasks produce varied results when fed with different inputs that only differ in terms of perceptually or musically inconsequential changes \TODO{Genre + deep net ref}. Whether intelligent machines should learn to perform a computational task similar to how humans do or not is a topic for debate \TODO{ref}. Irrespective of that, the approaches in \gls{mir} should at least scale to the repertoire on which they are developed and should be invariant to trivial low-level changes. The arguments are if these different inputs with trivial differences are fed to the system during the training phase, they can be well handled and learned. Probably true to an extent, but in \gls{mir} its virtually impossible to work with the whole real-world collection of music and most of the test datasets tend to be a small fraction of that. Therefore, robustness to trivial changes might be important to scale the approaches and systems developed in \gls{mir}. Moreover, one of the motivations of several researchers behind developing such systems is to learn more and get novel insights into music and validate our understanding of music, which might be futile if the systems are not really learning the underlying musical concepts. In this dissertation we try to steer away from such pitfalls by laying emphasis on the human interpretability of the research results, and wherever possible, of the output of the intermediate stages of our approaches. We are very aware that human interpretability doesn't guarantee that the approaches are generalizable or scalable, but, at least we are conscious of what our systems are learning, which eventually helps to improve them further. In addition, we work with a diverse, representative and sizable collection of \gls{iam} that contains editorial metadata and other relevant information in addition to the audio recordings, which can be regarded as a representative subset of the real-world audio collection of \gls{iam}~(\chapref{chap:corpus_music_corpora_and_datasets}).

%
%\begin{itemize}
%	\item What is music information retrieval
%	\item What is the context in which music information retrieval is growing
%	\item Why is MIR getting more important and finding many applications
%	\item Successful examples of MIR
%	\item What are the different contexts related with music that MIR can help in
%	\item what more?
%\end{itemize}

\subsection{CompMusic Project}
\label{sec:intro_motivation_compmusic}

Despite great advancements in the field of \gls{mir} in the last two decades, the outcome and generated knowledge to a large extent is not directly applicable to several music cultures of the world. This can be largely attributed to the fact that the research in \gls{mir} has been mainly focused on the western commercial music of the past few decades~\citep{XavierSerra2011}, and therefore, it does not extend to other music traditions. The western commercial music has shaped the undertaken research problems in \gls{mir}, and as a natural consequence, the obtain solutions are best suited for this music repertoire. These solutions often fail to respond to the multicultural reality present in the musics of the world~\citep{XavierSerra2011}. To gain insights on the factors responsible for capping the performance of the current \gls{mir} approaches (\secref{sec:intro_motivation_mir}) and to make the research outcome applicable to different music traditions of the world, development of these approaches in a multicultural context is important. Note that by this we do not imply that research  in \gls{mir} is only focused on western popular music. There has been work done on analysis of flamenco, folk, ethnomusicology, western classical \TODO{rephrase last line and put references}.

CompMusic\footnote{http://compmusic.upf.edu/} (Computational Models for the Discovery of the World'd Music) is a research project funded by the European Research Council that was born from the concerns mentioned above~\citep{XavierSerra2011}. The project focuses on five music traditions of the world: Hindustani (North India), Carnatic (South India), Turkish-makam (Turkey), Arab-Andalusian (Maghreb), and Beijing Opera (China). One of the main objectives of the CompMusic project is to promote and develop multicultural perspectives in \gls{mir}. In particular, the project aims to advance in the automatic description of music within \gls{mir} by identifying musically relevant problems coming from culture-specific contexts and by developing domain specific approaches to solve them. The solutions devised in this project can result in new computational methodologies of interest for a wide variety of music information processing problems. Thus, addressing the research problems in the context of diverse music cultures will not only help in advancing the knowledge in the specific cultures, but also expand the scope of current research in \gls{mir}. Furthermore, it can help bridge the semantic-gap and push the glass-ceiling. \TODO{Something on stress on good data, and how can it improve the systems}

The work presented in this dissertation for description and characterization of melodies in \gls{iam} has been carried out as a part of the CompMusic project and it aligns with the goals of the project. The cultural specificities of \gls{iam} have shaped our research work at each step ranging from the identification of research problems and building the data corpus to the choices made for different procedures and parameter settings. The insights gained during the handling of the challenges posed by the peculiar characteristics of melodies in \gls{iam} (\secref{sec:intro_challenges_oppurtunities}) will help expand the scope of the existing approaches in \gls{mir}. Our work paves the way for cross-cultural studies, which can help us get better insights into the influence of cultural training on perception and cognition of different musical aspects.


%\begin{itemize}
%	\item MIR is important and things are going fine, what's the issue then? Is there any bias in the kind of music tradition being analyzed in MIR right now? whats the reason for this bias
%	\item What kind of music traditions are kind of ignored for anlaysis now?
%	\item What's the reason why no one wants to consider them?
%	\item Is it even relevant to consider them? Which ones amongst them are the easiest ones or the most relevant ones to consider
%	\item Does everyone get benefitted from this kind of work? how things improve overall by doing this?
%	\item What is the objective of COmpMusic project, which music traditions it focuses on. 
%	\item What kind of methodologies are being worked upon in compmusic, What is the main philosophy
%	\item The idea of open-acess and reproducibility
%	\item what more?
%\end{itemize}

\subsection{Description and Characterization of Melodies}
\label{sec:intro_description_characterization_melodies}

\blockquote[{\cite{ringer2001melody}}]{``Melody, defined as pitched sounds arranged in musical time in accordance with given cultural conventions and constraints, represents a universal human phenomenon traceable to prehistoric times...wherever music became an intrinsic condition of life, certain common melodic procedures were necessarily adopted, because they satisfied basic physiological and biological requirements, if not the aesthetic imagination, per se....`Melodic unity' is configurationally speaking an intrinsic psychoacoustical function of melodic generation in a given historical-cultural context and music in the end be experienced as such...Above all, as a social product, melody is part and parcel of the cultural or subculture to which it owes its existence..."}

Essentially, there are two points to note here in the context of this dissertation; first, that melody is innate to humans and is one of the fundamental aspects in most music traditions, and second, the perception of melody is deeply effected by cultural influences. The innate characteristics of melody in human life is evident from a number of activities that surround us,  for example, lullabies, nursery rhythms, importance of intonation in languages, and in music. Melody also facilitate memorization. Even historically, when the printed medium was not prevalent, knowledge was orally passed on over generations, often set to music to make memorization easier. Even today, as listeners, very often we try to reproduce the melody of the song we want to remember. The importance of melody in our musical experiences makes its analysis an interesting topic for scientific study. With no surprises, a significant part of the research efforts in \gls{mir} is dedicated to computational analysis of melodic aspects of music.





As described by XXX "Melody is". Melody is also said imbibed in human XXX, basically when a child cries, intonatino in languages etc etc. Description and characterization of melodies is thus a fundamental task in computational analysis of music in \gls{mir}. Such a description can enable XXX, XX ,XX. 

Melodies can also be seen as being structured hierarchically according to the melodic framework prevalent in a music tradition. Starting from the basic unit, which is a pitched sound, a note, through a temporal interplay of such events building the whole melody. TYpically arrangement is such that notes group together into phrases, which is a unit that can be considered as a unit incorporating artists thoughts. Silence plays an important role in this structrures and a delimiter. Patterns play an important role in comprehension of melodies and are very important to be studied. It is also linked to lyrics and the strucrture in the lyrics, different lines, stanjas and probably like verses, chorus, which is another hrirarchy of this structure. And then when seen as a whole the fullpiece is basically a piece of art conveying what artist wanted to express and musically is making a unit of a framework (like raga tala). Issue of similarity

Description and characterization are umbrella terms involving a number of computational tasks that act at different levels of abstraction and musical concepts.  Study and analysis of each of these units and different strucrtures at different level is contributing towards description and characterization of melodies. Studing XX is helpgul in XXX and form the basis or input for other tasks. Studying XXX is helpful in blah blah.



Several different ways to approach description and characterization of music. one of whch is through doing the same for the melody. very widely addressed..
Automatic analysis for description and characterization of melodies is an imp task, core in mir. examples of the tasks that do that...

Melodies is composed of xxxx, heirarchical arrangement. Tonal and Temporal arramgement. 

One of the ways is pattenr, very imp. 

We can describe at diff level, notes it uses, key and mode, stylistic description, structural description, description of the characterisics

In this thesis focus on patterns. 

what is raga and makam commonly called asd doing that is imp. and why


so basically melody is imp, melodic desc is imp, it can help in XXXXXX, it can be at various levels, low level to high level, each one has its own imp, we focus more on patterns, patterns are imp, and why so..then melodic framework, the most abstracted higher level view, we do that too, and since that is closely linked with mood and emotions that directly becomes relevant for recommendation and navigation and those kind of stuff

description 
\section{Melodic Analysis in Indian Art Music: Challenges and Opportunities}
\label{sec:intro_challenges_oppurtunities}

%
%\begin{itemize}
%	\item in general music similaity, search and discovery. Different applications and context
%	\item melodic analysis->representation of tonal content often used->melodic representaiton->difficulaty in extraction
%	\item problems which are solved by the state of the art in melody extraction for different music types
%	\item for which music types melodic anlaysis make more sense and have been done successfully.
%	\item Symbolic domain research, a lot of work.
%	\item Key detection and cover song detection.
%\end{itemize}
%
%\begin{itemize}
%	\item Music similarity focus in MIR
%	\item application and context of music similarity
%	\item how that has given rise to search and discovery in different contexts
%	\item use cases/ applications / relevant problems within similarity and search and discovery
%\end{itemize}

\section{Scope and Objectives}
\label{sec:intro_scope_context_relevance}

we take an engineering perspective, MIR types approach of describing music, within and following CompMusic ideologies, culture specific in a multicultral context, applied, and micture of culture specific and 

In this section include both broad objectives of this thesis and present a list of detailed research tasks done, since they are quite many in numbers that will give an idea about how they relate to the overall goal and this is the only place where we are going to enurate them.

\section{Organization and Outline of the thesis}
\label{sec:intro_organization}
